%==============================================================================
\documentclass[MathematicsNumericsDerivationsAndOpenFOAM.tex]{subfiles}
\begin{document}
%==============================================================================


\pagenumbering{roman}
\newpage
%
%
    \textbf{T}he following book is related to the \OF version maintained and
    distributed by the OpenFOAM Foundation.

%
%
\vspace{20pt}
%
%
    \textbf{T}he c++ code snippets presented in this book are related to
    the \OF commit \DEV.

%
%
\vspace{20pt}
%
%
    \textbf{T}here is no warranty that the content of this book is errorless.
    Tobias Holzmann tried to incorporate his knowledge and the latest
    developments into this book. Nevertheless, typos and/or wrong code snippets
    cannot be excluded. If you -- \textit{the community} -- find any mistake,
    please don't hesitate to contact Tobias Holzmann via Book@Holzmann-cfd.com.

    This however, presumes that you read this part of the book and \textit{reading}
    seems to be wired today.
%
%
\newpage
%
%
%
\textit{
    \textbf{T}his book is dedicated to the \OF community and to all colleagues
    and people who are interested in computational fluid dynamics, numerical
    simulations and who supports Tobias' work he shares on his website Holzmann
    CFD. Tobias' ambition to write the book was based on his personal love to
    the open-source thought. Thus, his goal and wish are to give any CFD lover
    an introduction to computational fluid dynamics while showing interesting
    equations and exciting relations that are not obviously presented in most
    of the famous books and papers in that particular area. Additionally, the
    book should prepare everybody for the tasks that one may work on during
    her/his personal career, hopefully with \OF. Tobias wishes everybody all the
    best and an exciting time with \OF and the beautilful subject named CFD.
}

%
%
\vspace{20pt}
%
%
\textit{
    \textbf{T}he book is available as an free and open-source version and a full 
    version. If one is interested in a hard-cover, do not hesitate to write an email
    to \href{mailto:book@Holzmann-cfd.com}{\color{myCiteColor}book@Holzmann-cfd.com}.
}

%
%
\vspace{20pt}
%
%
\textit{
    \textbf{B}e aware the this book is updated more often. The \OF community do
    have access to the latest corrections of all parts that are available for free.
    All those who do have the full version of the book will always have access to
    all corrections and features except if the new version includes a lot of new
    material. However, you will then have the possibility to get the extension
    with a discount of 80 \%.
}

%
%
\vspace{20pt}
%
%
\textit{
    \textbf{T}o get further information about projects, developments and the
    latest news of Tobias work --- primarily for the community ---, feel free to
    follow Tobias Holzmann on \texttt{Twitter, LinkedIn, XING or Youtube}.
}

%
%
\vspace{20pt}
%
%
\textit{
    \textbf{E}ach feedback is warmly welcomed and will be considered in new
    releases. If you find mistakes, typos or if you want to add a chapter,
    feel free to contact Tobias \href{mailto:community@Holzmann-cfd.com}{\color{myCiteColor}community@Holzmann-cfd.com}
    or use the public repository, extend it and make a pull request.
}

%
%
\vspace{20pt}
%
%
\textit{
    \textbf{C}orperations are welcomed. If one wants to contribute to the book,
    you can use the public repository, extend the latex code and add your
    code to the repository. Tobias would feel honored if one is extending it.
    If you want to translate the book into your language, feel free to create
    a new branch, and translate it.
}

%
%
\vspace{20pt}
%
%
\textit{
    \textbf{S}upport Tobias work with a voluntary donation. Writing the book
    was a time consuming time for Tobias and it offers a lot of information,
    tips and tricks as well as some interesting not-always-obvious correlations.
    \url{https://Holzmann-cfd.com/}.
}

%
%
\newpage


%==============================================================================
\end{document}
%==============================================================================
